---
title: "Cox Model"
output: html_document
date: "2023-12-18"
---

```{r}
library(readxl)
data <- read_excel("../dataset/TrainingSet.xlsx", col_names = TRUE)
```

Libraries:
```{r,'error=FALSE', 'warning=FALSE'}
library(readxl)
library(survival)
library(survminer)
library(dplyr) 
library(ggplot2)
library(knitr)
library(broom)
library(tidyr)
```

Summary:
```{r}
summary(data)
```
Conto numero eventi per ogni covariata
```{r}
table(data$Users, data$Status)
table(data$Purpose, data$Status)
table(data$Orbit, data$Status)
table(data$Continent, data$Status)


```
Dalla letteratura risulta che non ho abbastanza eventi per Navigation, Elliptical, Africa, Multinational, Oceania
Simulation work has suggested that at least 10 events need to be observed for each covariate considered, and anything less will lead to problems.
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2376927/

Rimuoviamo queste classi
```{r}
i <- which(data$Purpose == 'Navigation' | data$Orbit == 'Elliptical' | data$Continent == 'Africa' | data$Continent == 'Multinational' | data$Continent == 'Oceania' )
data <- data[-i,]
```

A livello di continenti è un confronto tra America, Asia e Europa: Maggiori potenze economiche 

Controlliamo
```{r}
table(data$Users, data$Status)
table(data$Purpose, data$Status)
table(data$Orbit, data$Status)
table(data$Continent, data$Status)
table(data$Status)
```


```{r}
time <- data$`Effective Lifetime`
status <- data$Status

covariates <- c( "Perigee" ,"Apogee", 'Eccentricity', "Inclination", "Period", "Mass", "Users", "Purpose", "Continent", "Orbit")

cox1 <- coxph(Surv(time,status == 'retired') ~ ., data = data[,covariates])
summary(cox1)

```

Strata su Purpose
```{r}
time <- data$`Effective Lifetime`
status <- data$Status

cox2 <- coxph(Surv(time,status == 'retired') ~ Perigee + Apogee + Eccentricity + Inclination + Period +
                Mass + Users + strata(Purpose) + Continent + Orbit, data = data)
summary(cox2)

```


Verifico Assunzioni
```{r}
test <- cox.zph(cox2)
test
```

Strata su Orbit
```{r}
time <- data$`Effective Lifetime`
status <- data$Status

cox3<- coxph(Surv(time,status == 'retired') ~ Perigee + Apogee + Eccentricity + Inclination + Period +
                Mass + Users + strata(Purpose) + Continent + strata(Orbit), data = data)
summary(cox3)

```


Rimuovo Perigeo
```{r}
time <- data$`Effective Lifetime`
status <- data$Status

cox4<- coxph(Surv(time,status == 'retired') ~  Apogee + Eccentricity + Inclination + Period +
                Mass + Users + strata(Purpose) + Continent + strata(Orbit), data = data)
summary(cox4)

```

Rimuovo Periodo

```{r}
time <- data$`Effective Lifetime`
status <- data$Status

cox5<- coxph(Surv(time,status == 'retired') ~  Apogee + Eccentricity + Inclination  +
                Mass + Users + strata(Purpose) + Continent + strata(Orbit), data = data)
summary(cox5)

```

Rimuovo Inclination

```{r}
time <- data$`Effective Lifetime`
status <- data$Status

cox6<- coxph(Surv(time,status == 'retired') ~  Apogee + Eccentricity   +
                Mass + Users + strata(Purpose) + Continent + strata(Orbit), data = data)
summary(cox6)

```
Concordance: The concordance statistic C is defined as the fraction of
concordant pairs, and is an estimate of P(xi > xj |yi > yj ), or equivalently P(yi > yj |xi > xj ).
The first form might make more sense if, for instance, one took a list of elected officials, say, and
looked back at local newspapers to see how many of the election contests had been predicted
correctly (probablility the prediction is correct, given outcome).

Verifico Assunzioni
```{r}
test <- cox.zph(cox6)
test
```
Non sono super verificate per Apogeo ma ce ne facciamo una ragione...

```{r}
ggcoxdiagnostics(cox6, type = "scaledsch")
```

Martingale residuals 
```{r}
ggcoxdiagnostics(cox6, type = "martingale") # Diffcili da interpretare

```
Deviance Residuals
Deviance Residuals:
These residuals should be roughly symmetrically distributed about zero
with a standard deviation of 1.

-   Positive values correspond to individuals that "died too soon"
    compared to expected survival times.
-   Negative values correspond to individual that "lived too long".
-   Very large or small values are outliers, which are poorly predicted
    by the model.
```{r}
ggcoxdiagnostics(cox6, type = "deviance")

```
calcolo AIC
```{r}
AIC.cox6 = AIC(cox6)
```

## Provo a trasformare Eccentricità

Certainly, in certain scientific and engineering contexts, the logarithm of the eccentricity of an orbit is often used. In celestial mechanics, particularly when dealing with orbits, the eccentricity of an ellipse is a parameter that describes how much the orbit deviates from being a perfect circle.

Trasformo Eccentricità per avere un miglior grafico dei residui
```{r}
i <- which(data$Eccentricity == 0)
data$Eccentricity[i] = 1
data$log.Eccentricity <- -log(data$Eccentricity)
```


```{r}
time <- data$`Effective Lifetime`
status <- data$Status

covariates <- c( "Perigee" ,"Apogee", 'log.Eccentricity', "Inclination", "Period", "Mass", "Users", "Purpose", "Continent", "Orbit")

cox1bis <- coxph(Surv(time,status == 'retired') ~ ., data = data[,covariates])
summary(cox1bis)

```

Rimuovo Perigee
```{r}
time <- data$`Effective Lifetime`
status <- data$Status

covariates <- c("Apogee", 'log.Eccentricity', "Inclination", "Period", "Mass", "Users", "Purpose", "Continent", "Orbit")

cox2bis <- coxph(Surv(time,status == 'retired') ~ ., data = data[,covariates])
summary(cox2bis)

```

Strata su purpose
```{r}
time <- data$`Effective Lifetime`
status <- data$Status

cox2bis <- coxph(Surv(time,status == 'retired') ~  Apogee + log.Eccentricity + Inclination + Period +
                Mass + Users + strata(Purpose) + Continent + Orbit, data = data)
summary(cox2bis)

```

Rimuovo Inclination
```{r}
time <- data$`Effective Lifetime`
status <- data$Status

cox3bis <- coxph(Surv(time,status == 'retired') ~  Apogee + log.Eccentricity  + Period +
                Mass + Users + strata(Purpose) + Continent + Orbit, data = data)
summary(cox3bis)

```

Verifico Assunzioni
```{r}
test <- cox.zph(cox3bis)
test
```

Strata su orbit

```{r}
time <- data$`Effective Lifetime`
status <- data$Status

cox4bis <- coxph(Surv(time,status == 'retired') ~  Apogee + log.Eccentricity  + Period +
                Mass + Users + strata(Purpose) + Continent + strata(Orbit), data = data)
summary(cox4bis)

```


Rimuovo Periodo

```{r}
time <- data$`Effective Lifetime`
status <- data$Status

cox5bis <- coxph(Surv(time,status == 'retired') ~  Apogee + log.Eccentricity   +
                Mass + Users + strata(Purpose) + Continent + strata(Orbit), data = data)
summary(cox5bis)

```

Verifico Assunzioni
```{r}
test <- cox.zph(cox5bis)
test
```
Tutte le assunzioni sono verificate
Plot dei residui

```{r}
ggcoxdiagnostics(cox5bis, type = "scaledsch")
```

Martingale residuals Difficili da interpretare
```{r}
ggcoxdiagnostics(cox5bis, type = "martingale")

```
Deviance Residuals:
These residuals should be roughly symmetrically distributed about zero
with a standard deviation of 1.

-   Positive values correspond to individuals that "died too soon"
    compared to expected survival times.
-   Negative values correspond to individual that "lived too long".
-   Very large or small values are outliers, which are poorly predicted
    by the model.
    
Direi ok!
```{r}
ggcoxdiagnostics(cox5bis, type = "deviance")

```

calcolo AIC
```{r}
AIC.cox5bis = AIC(cox5bis)
```

Guardando l'AIC è meglio trasformare l'eccentricità.

## Previsione

```{r}
library(coxed)

ed1 <- coxed(cox5bis, method="npsf")
predlife=ed1$exp.dur
#mean and median of the predicted durations:
ed1$mean
ed1$median

# PLOT: The estimated cumulative baseline hazard function and survivor function
baseline <- gather(ed1$baseline.functions, cbh, survivor, key="survivefunction", value="value")
ggplot(baseline, aes(x=time, y=value)) +
     geom_line() +
     xlab("Time") +
     ylab("Function") +
     facet_wrap( ~ survivefunction, scales = "free")
```
PREDICTION

Preparazione test set
```{r}
test.set <- read_excel("../Dataset/TestSet.xlsx", col_names = TRUE)
```


```{r}
i <- which(test.set$Purpose == 'Navigation' | test.set$Orbit == 'Elliptical' | test.set$Continent == 'Africa' | test.set$Continent == 'Multinational' | test.set$Continent == 'Oceania' )
test.set <- test.set[-i,]
```

```{r}
table(test.set$Users, test.set$Status)
table(test.set$Purpose, test.set$Status)
table(test.set$Orbit, test.set$Status)
table(test.set$Continent, test.set$Status)
table(test.set$Status)
```
Trasformo Eccentricità per avere un miglior grafico dei residui
```{r}
i <- which(test.set$Eccentricity == 0)
test.set$Eccentricity[i] = 1
test.set$log.Eccentricity <- -log(test.set$Eccentricity)
```



Per ottenere baseline hazard dei nuovi dati
```{r}
bashaz.train <- basehaz(cox5bis)
bashaz.train
table(bashaz.train$strata) # strata su orbit e Purpose --> 8 combinazioni in teoria
```


```{r}
bashaz.test <- basehaz(cox5bis, test.set, centered=TRUE)
bashaz.test
table(bashaz.test$strata)
# technology DEvelopment, Geo non esiste  
table(test.set$Orbit, test.set$Purpose)
# technology DEvelopment, Geo non esiste neanche nel test set -> buono 

```
strata = 1 --> hazard riferita alla prima osservazione


```{r}
## S3 method for class 'coxph'
expLP = predict(cox5bis, test.set,
#type=c("lp", "risk", "expected", "terms", "survival"),
type = 'risk',
reference=c("strata", "sample", "zero"))
```

Prediction
```{r}
# a mano: Idea imprlementare quello che fa Coxed (Incluso bootstrap) in modo da poter usare strata
# Qui il link di coxed : https://github.com/jkropko/coxed/blob/master/R/coxed.R
```


```{r}
coxed <- coxed::coxed(cox5bis, newdata = test.set, method="npsf", bootstrap = TRUE, B=750)
predicted_lifetime <- coxed$exp.dur
```



```{r}
id <- which(test.set$Status == 'retired')
diff <- predicted_lifetime$exp.dur[id] - test.set$`Effective Lifetime`[id]
plot(diff)

```
```{r}
sqrt(sum(diff^2)/length(diff))
mean(diff)
median(diff)
#mean(test.set$`Effective Lifetime`[id])
```
```{r}


# Calculate Concordance Index
c_index <- survConcordance(Surv(test.set$`Effective Lifetime`, predicted_lifetime$exp.dur) ~ predicted_lifetime$exp.dur)
print(paste("Concordance Index: ", c_index$concordance))

```
```{r}
diff <- predicted_lifetime$exp.dur[id] - test.set$`Expected Lifetime`[id]
plot(diff)
```

```{r}
sorted_lt <- predicted_lifetime[order(predicted_lifetime$exp.dur),]

t <- seq(0, max(test.set$`Effective Lifetime`), length = length(sorted_lt$exp.dur))
plot(t, sorted_lt$exp.dur, type = 'l', lwd = 2)
lines(t, sorted_lt$lb)
lines(t, sorted_lt$ub)
```

