---
title: "NPS Project Cox Model - discretize var"
output: html_document
date: "2023-12-05"
---


```{r}
library(readxl)
data <- read_excel("../Dataset/JoinDatasets.xlsx", col_names = TRUE)
```

Libraries:
```{r,'error=FALSE', 'warning=FALSE'}
library(readxl)
library(survival)
library(survminer)
library(dplyr) 
library(ggplot2)
library(knitr)
library(broom)
library(tidyr)
```


Summary:
```{r}
summary(data)
```

Rendi colonne factor:
```{r}
data$Country <- factor(data$Country, ordered = F)
print('Country')
table(data$Country)

data$Continent <- factor(data$Continent, ordered = F)
print('Continent')
table(data$Continent)

data$Users <- factor(data$Users, ordered = F)
print('Users')
table(data$Users)

data$Purpose <- factor(data$Purpose, ordered = F)
print('Purpose')
table(data$Purpose)

data$Orbit <- factor(data$Orbit, ordered = F)
print('Class of Orbit')
table(data$Orbit)

data$Status <- factor(data$Status, ordered = F)
print('Status')
table(data$Status)
```

Mischia righe
```{r}
data <- data[sample(nrow(data)), ]
```

Creo dummy
Creo variabili dummy
```{r}
#install.packages("fastDummies")
library(fastDummies)

# USERS : Levels: Civil Commercial Government Military

dummy_data <- dummy_cols(data, select_columns = "Users", remove_first_dummy = TRUE)

# PURPOSE : Levels: Communications Earth Observation Navigation Space Science Technology Development

dummy_data <- dummy_cols(dummy_data, select_columns = "Purpose", remove_first_dummy = TRUE)


# ORBIT : Levels: Elliptical GEO LEO MEO

dummy_data <- dummy_cols(dummy_data, select_columns = "Orbit", remove_first_dummy = TRUE)


# CONTINENT: Levels: Africa Americas Asia Europe Multinational Oceania

dummy_data <- dummy_cols(dummy_data, select_columns = "Continent", remove_first_dummy = TRUE)
```


Exploration of numerical variables:
```{r}
num <- c("Longitude of GEO", "Perigee", "Apogee", "Eccentricity", "Inclination", "Period", "Mass")
data_num <- data[,num]

cor(data_num)
# Apogee - Period 0.96672436  
# Perigee - Inclination -0.88628415   
# Perigee - Period 0.8327784  
# Period - Inclination -0.73233046  
# Apogee - Perigee 0.67998408  
# Apogee - Eccentricity 0.67998408  
# Apogee - Inclination -0.60232423  

pairs(data_num)
```
Mass: include it
```{r}
plot(data$Mass, data$`Effective Lifetime`, xlab='Mass')
hist(data$Mass, xlab='Mass', main='Histogram of Mass', prob = TRUE)

summary(data$Mass)
```

Discretiziamo la massa

```{r}
Massa.disc <- cut(data$Mass, breaks=c(0, 1000, Inf), labels=c("light", "heavy"))
dummy_data$Mass.disc <- Massa.disc

```


Period:
```{r}
plot(data$Period, data$`Effective Lifetime`, xlab='Period')
hist(data$Period, xlab='Period', main='Histogram of Period', prob = TRUE)

summary(data$Period)
```

Discretiziamo il periodo

```{r}
Per.disc <- cut(data$Period, breaks=c(0, 500, 1000, Inf), labels=c("short", "medium", "long"))
dummy_data$Per.disc <- Per.disc

```


Inclination: 
```{r}
plot(data$Inclination, data$`Effective Lifetime`, xlab='Inclination')
hist(data$Inclination, xlab='Inclination', main='Histogram of Inclination', prob = TRUE)

summary(data$Inclination)
```
Discretiziamo l'inclination

```{r}
Inc.disc <- cut(data$Inclination, breaks=c(-Inf, 40, 70, Inf), labels=c("low", "medium", "high"))
dummy_data$Inc.disc <- Inc.disc

```


Eccentricity: ( Non faccio nulla, non è significativa)
```{r}
plot(data$Eccentricity, data$`Effective Lifetime`, xlab='Eccentricity')
hist(data$Eccentricity, xlab='Eccentricity', main='Histogram of Eccentricity', prob = TRUE)

summary(data$Eccentricity)
```
Apogee
```{r}
plot(data$Apogee, data$`Effective Lifetime`, xlab='Apogee')
hist(data$Apogee, xlab='Apogee', main='Histogram of Apogee', prob = TRUE)

summary(data$Apogee)
```
Discretiziamo Apogeo

```{r}
Ap.disc <- cut(data$Apogee, breaks=c(0, 25000, Inf), labels=c("low", "high"))
dummy_data$Ap.disc <- Ap.disc

```


Perigee
```{r}
plot(data$Perigee, data$`Effective Lifetime`, xlab='Perigee')
hist(data$Perigee, xlab='Perigee', main='Histogram of Perigee', prob = TRUE)

summary(data$Perigee)
```

Discretiziamo Perigee

```{r}
Pe.disc <- cut(data$Perigee, breaks=c(0, 5000, 25000, Inf), labels=c("low","medium","high"))
dummy_data$Pe.disc <- Pe.disc

```


Longitude of GEO:
```{r}
plot(data$`Longitude of GEO`, data$`Effective Lifetime`, xlab='`Longitude of GEO`')
hist(data$`Longitude of GEO`, xlab='`Longitude of GEO`', main='Histogram of `Longitude of GEO`', prob = TRUE)

summary(data$`Longitude of GEO`)
```

Creo variabili dummy
```{r}

# MASS: Levels: light heavy

dummy_data <- dummy_cols(dummy_data, select_columns = "Mass.disc", remove_first_dummy = TRUE)

# PERIOD: Levels: short medium long 

dummy_data <- dummy_cols(dummy_data, select_columns = "Per.disc", remove_first_dummy = TRUE)

# INCLINATION: Levels: low medium high 

dummy_data <- dummy_cols(dummy_data, select_columns = "Inc.disc", remove_first_dummy = TRUE)

# PERIGEE: Levels: low medium high  

dummy_data <- dummy_cols(dummy_data, select_columns = "Pe.disc", remove_first_dummy = TRUE)

# APOGEE: Levels: low high

dummy_data <- dummy_cols(dummy_data, select_columns = "Ap.disc", remove_first_dummy = TRUE)
```


```{r}
# dataset senza Purpose, Users, Continent, Orbit, Mass, Period,  Inclination, Apogee and Perigee

dummy_data <- dummy_data[,c(1:2,6,9,12:16,19:33, 39:46)]

```

### COX MODEL
```{r}
time <- dummy_data$`Effective Lifetime`
status <- dummy_data$Status


covariates <-
  c(
    "Longitude of GEO" ,
    "Eccentricity" ,
    "Users_Commercial",
    "Users_Government",
    "Users_Military",
    "Purpose_Earth Observation",
    "Purpose_Navigation",
    "Purpose_Space Science",
    "Purpose_Technology Development",
    "Orbit_GEO",
    "Orbit_LEO",
    "Orbit_MEO",
    "Continent_Americas",
    "Continent_Asia",
    "Continent_Europe",
    "Continent_Multinational",
    "Continent_Oceania",
    "Mass.disc_heavy",
    "Per.disc_medium",
    "Per.disc_long",
    "Inc.disc_medium",
    "Inc.disc_high",
    "Pe.disc_medium",
    "Pe.disc_high",
    "Ap.disc_high"
    
  )

n_coef <- length(covariates)

cox1 <- coxph(Surv(time,status == 'retired') ~ ., data = dummy_data[,covariates])
summary(cox1)
```

```{r}
time <- dummy_data$`Effective Lifetime`
status <- dummy_data$Status


covariates <-
  c(
    "Longitude of GEO" ,
    "Eccentricity" ,
    "Users_Commercial",
    "Users_Government",
    "Users_Military",
    "Purpose_Earth Observation",
    "Purpose_Navigation",
    "Purpose_Space Science",
    "Purpose_Technology Development",
    "Orbit_GEO",
    "Orbit_LEO",
    "Orbit_MEO",
    "Continent_Americas",
    "Continent_Asia",
    "Continent_Europe",
    "Continent_Multinational",
    "Continent_Oceania",
    "Mass.disc_heavy",
    "Per.disc_medium",
    "Per.disc_long",
    "Inc.disc_medium",
    "Inc.disc_high",
    "Pe.disc_medium",
    "Ap.disc_high"
    
  )

n_coef <- length(covariates)

cox2 <- coxph(Surv(time,status == 'retired') ~ ., data = dummy_data[,covariates])
summary(cox2)
```

```{r}
time <- dummy_data$`Effective Lifetime`
status <- dummy_data$Status


covariates <-
  c(
    "Longitude of GEO" ,
    "Eccentricity" ,
    "Users_Commercial",
    "Users_Government",
    "Users_Military",
    "Purpose_Earth Observation",
    "Purpose_Navigation",
    "Purpose_Space Science",
    "Purpose_Technology Development",
    "Orbit_GEO",
    "Orbit_LEO",
    "Continent_Americas",
    "Continent_Asia",
    "Continent_Europe",
    "Continent_Multinational",
    "Continent_Oceania",
    "Mass.disc_heavy",
    "Per.disc_medium",
    "Per.disc_long",
    "Inc.disc_medium",
    "Inc.disc_high",
    "Pe.disc_medium",
    "Ap.disc_high"
    
  )

n_coef <- length(covariates)

cox2 <- coxph(Surv(time,status == 'retired') ~ ., data = dummy_data[,covariates])
summary(cox2)
```

Remove Purpose_Navigation
```{r}
time <- dummy_data$`Effective Lifetime`
status <- dummy_data$Status


covariates <-
  c(
    "Longitude of GEO" ,
    "Eccentricity" ,
    "Users_Commercial",
    "Users_Government",
    "Users_Military",
    "Purpose_Earth Observation",
    "Purpose_Navigation",
    "Purpose_Space Science",
    "Purpose_Technology Development",
    "Orbit_GEO",
    "Orbit_LEO",
    "Orbit_MEO",
    "Continent_Americas",
    "Continent_Asia",
    "Continent_Europe",
    "Continent_Multinational",
    "Continent_Oceania",
    "Mass.disc_heavy",
    "Per.disc_medium",
    "Per.disc_long",
    "Inc.disc_medium",
    "Inc.disc_high",
    "Pe.disc_medium",
    "Ap.disc_high"
    
  )

n_coef <- length(covariates)

cox2 <- coxph(Surv(time,status == 'retired') ~ ., data = dummy_data[,covariates])
summary(cox2)
```

```{r}
time <- dummy_data$`Effective Lifetime`
status <- dummy_data$Status


covariates <-
  c(
    "Longitude of GEO" ,
    "Eccentricity" ,
    "Users_Commercial",
    "Users_Government",
    "Users_Military",
    "Purpose_Earth Observation",
    "Purpose_Space Science",
    "Purpose_Technology Development",
    "Orbit_GEO",
    "Orbit_LEO",
    "Continent_Americas",
    "Continent_Asia",
    "Continent_Europe",
    "Continent_Multinational",
    "Continent_Oceania",
    "Mass.disc_heavy",
    "Per.disc_medium",
    "Per.disc_long",
    "Inc.disc_medium",
    "Inc.disc_high",
    "Pe.disc_medium",
    "Ap.disc_high"
    
  )

n_coef <- length(covariates)

cox3 <- coxph(Surv(time,status == 'retired') ~ ., data = dummy_data[,covariates])
summary(cox3)
```
Remove Users_Military

```{r}
time <- dummy_data$`Effective Lifetime`
status <- dummy_data$Status


covariates <-
  c(
    "Longitude of GEO" ,
    "Eccentricity" ,
    "Users_Commercial",
    "Users_Government",
    "Purpose_Earth Observation",
    "Purpose_Space Science",
    "Purpose_Technology Development",
    "Orbit_GEO",
    "Orbit_LEO",
    "Continent_Americas",
    "Continent_Asia",
    "Continent_Europe",
    "Continent_Multinational",
    "Continent_Oceania",
    "Mass.disc_heavy",
    "Per.disc_medium",
    "Per.disc_long",
    "Inc.disc_medium",
    "Inc.disc_high",
    "Pe.disc_medium",
    "Ap.disc_high"
    
  )

n_coef <- length(covariates)

cox4 <- coxph(Surv(time,status == 'retired') ~ ., data = dummy_data[,covariates])
summary(cox4)
```
Remove Inc.disch_high
```{r}
time <- dummy_data$`Effective Lifetime`
status <- dummy_data$Status


covariates <-
  c(
    "Longitude of GEO" ,
    "Eccentricity" ,
    "Users_Commercial",
    "Users_Government",
    "Purpose_Earth Observation",
    "Purpose_Space Science",
    "Purpose_Technology Development",
    "Orbit_GEO",
    "Orbit_LEO",
    "Continent_Americas",
    "Continent_Asia",
    "Continent_Europe",
    "Continent_Multinational",
    "Continent_Oceania",
    "Mass.disc_heavy",
    "Per.disc_medium",
    "Per.disc_long",
    "Inc.disc_medium",
    "Pe.disc_medium",
    "Ap.disc_high"
    
  )

n_coef <- length(covariates)

cox5 <- coxph(Surv(time,status == 'retired') ~ ., data = dummy_data[,covariates])
summary(cox5)
```

Rmevove Inc.disc_medium
```{r}
time <- dummy_data$`Effective Lifetime`
status <- dummy_data$Status


covariates <-
  c(
    "Longitude of GEO" ,
    "Eccentricity" ,
    "Users_Commercial",
    "Users_Government",
    "Purpose_Earth Observation",
    "Purpose_Space Science",
    "Purpose_Technology Development",
    "Orbit_GEO",
    "Orbit_LEO",
    "Continent_Americas",
    "Continent_Asia",
    "Continent_Europe",
    "Continent_Multinational",
    "Continent_Oceania",
    "Mass.disc_heavy",
    "Per.disc_medium",
    "Per.disc_long",
    "Pe.disc_medium",
    "Ap.disc_high"
    
  )

n_coef <- length(covariates)

cox6 <- coxph(Surv(time,status == 'retired') ~ ., data = dummy_data[,covariates])
summary(cox6)
```
Rmevove Purpose_Earth_obs
```{r}
time <- dummy_data$`Effective Lifetime`
status <- dummy_data$Status


covariates <-
  c(
    "Longitude of GEO" ,
    "Eccentricity" ,
    "Users_Commercial",
    "Users_Government",
    "Purpose_Space Science",
    "Purpose_Technology Development",
    "Orbit_GEO",
    "Orbit_LEO",
    "Continent_Americas",
    "Continent_Asia",
    "Continent_Europe",
    "Continent_Multinational",
    "Continent_Oceania",
    "Mass.disc_heavy",
    "Per.disc_medium",
    "Per.disc_long",
    "Pe.disc_medium",
    "Ap.disc_high"
    
  )

n_coef <- length(covariates)

cox7 <- coxph(Surv(time,status == 'retired') ~ ., data = dummy_data[,covariates])
summary(cox7)
```

Remove Purpose_Tech dev
```{r}
time <- dummy_data$`Effective Lifetime`
status <- dummy_data$Status


covariates <-
  c(
    "Longitude of GEO" ,
    "Eccentricity" ,
    "Users_Commercial",
    "Users_Government",
    "Purpose_Space Science",
    "Orbit_GEO",
    "Orbit_LEO",
    "Continent_Americas",
    "Continent_Asia",
    "Continent_Europe",
    "Continent_Multinational",
    "Continent_Oceania",
    "Mass.disc_heavy",
    "Per.disc_medium",
    "Per.disc_long",
    "Pe.disc_medium",
    "Ap.disc_high"
    
  )

n_coef <- length(covariates)

cox8 <- coxph(Surv(time,status == 'retired') ~ ., data = dummy_data[,covariates])
summary(cox8)
```

Remove Ap.disc high 
```{r}
time <- dummy_data$`Effective Lifetime`
status <- dummy_data$Status


covariates <-
  c(
    "Longitude of GEO" ,
    "Eccentricity" ,
    "Users_Commercial",
    "Users_Government",
    "Purpose_Space Science",
    "Orbit_GEO",
    "Orbit_LEO",
    "Continent_Americas",
    "Continent_Asia",
    "Continent_Europe",
    "Continent_Multinational",
    "Continent_Oceania",
    "Mass.disc_heavy",
    "Per.disc_medium",
    "Per.disc_long",
    "Pe.disc_medium"
    )

n_coef <- length(covariates)

cox9 <- coxph(Surv(time,status == 'retired') ~ ., data = dummy_data[,covariates])
summary(cox9)
```

Remove Orbit_LEO
```{r}
time <- dummy_data$`Effective Lifetime`
status <- dummy_data$Status


covariates <-
  c(
    "Longitude of GEO" ,
    "Eccentricity" ,
    "Users_Commercial",
    "Users_Government",
    "Purpose_Space Science",
    "Orbit_GEO",
    "Continent_Americas",
    "Continent_Asia",
    "Continent_Europe",
    "Continent_Multinational",
    "Continent_Oceania",
    "Mass.disc_heavy",
    "Per.disc_medium",
    "Per.disc_long",
    "Pe.disc_medium"
    )

n_coef <- length(covariates)

cox10 <- coxph(Surv(time,status == 'retired') ~ ., data = dummy_data[,covariates])
summary(cox10)
```

Verifichiamo assunzioni:

```{r}
test <- cox.zph(cox10)
test
```
Le assunzioni non sono verificate.



